### 2019-11-25 ic-journal/ccs.tex
text

Here we recall the core theory of CCS.
We assume a possibly infinite set of names $\mathscr{L} = \{a, b,
\ldots\}$ yielding input and $\overline{\mbox{output}}$ labels (actions), plus a special invisible
action $\tau \notin \mathscr{L}$, and another possibly infinite
set of agent variables \hl{$\mathscr{X} = \{X, Y, \ldots\}$}.
The class of \hl{all} CCS \hl{terms (or expressions)} is then inductively
defined from $\nil$ (the terminated process) and agent variables by the operators
of \emph{prefixing} (.), \emph{parallel composition} ($|$),
\emph{summation}
(or \emph{binary choice}, $+$),
\emph{restriction} ($\nu$), \emph{relabeling} ($[\cdot]$) and
\emph{recursion} ($\texttt{rec}$):
\begin{equation*}
\begin{array}{cccl}
\mu  & := & \tau\!\!\!\! & \midd \; a  \; \midd \;  \outC a  \\
P  & := & \nil\!\!\!\! & \midd \;  \mu . P \; \midd \;  P_1 |  P_2 \; \midd  \;
P_1 + P_2 \; \midd 
(\res L)\, P
\; \midd \; P\; [\mathit{rf}]  % relabelling
\;  \midd \;  X \; \midd \; \recu X P
\end{array}
\end{equation*}
In our presentation of CCS, the restriction operator takes a set of labels $L
  \subseteq \mathscr{L}$, rather than a single one.  
The relabeling operator takes a relabeling function
  $\mathit{rf} \colon \mathscr{L} \cup \overline{\mathscr{L}} \cup
    \{\tau\} \rightarrow \mathscr{L} \cup \overline{\mathscr{L}} \cup
    \{\tau\}$, that can  handle multiple actions including $\tau$. A valid
    relabeling function $\mathit{rf}$ must however 
satisfy $\mathit{rf}(\tau) =
    \tau$ and $\forall l\in\mathscr{L} \cup \overline{\mathscr{L}}.\, \mathit{rf}(\overline{l}) =
    \overline{\mathit{rf}(l)}$ (with $\bar{\bar l} = l$ for all $l \in
    \mathscr{L}$).
We sometimes omit a trailing $\nil$, e.g., writing $a|b$ for $a.\nil |b .\nil$.
A CCS process $P$ may evolve to another one, say $P'$, under
  an action $\mu$, written by $P \arr{\mu} P'$.
The transition semantics of CCS processes is given by means of
a Labeled Transition System (LTS) expressed in Structural Operational
Semantics (SOS) rules shown in Fig.~\ref{f:LTSCCS}.
(The symmetric versions of the SOS rules for
parallel composition and binary choice are omitted.)
A CCS processes uses only \emph{guarded sums} if all occurrences of
sums are in the form $a.P + b.Q$. 
The \emph{immediate derivatives} of a
process $P$ are elements of the set $\{P' \st P \arr\mu P' \mbox{for some $\mu$}\}$.
We use $\ell$ to range over visible actions (i.e.~inputs or outputs, excluding  $\tau$)
and $\alpha, \mu$  to range over all actions.
\begin{figure*}[t]
\begin{center}
\vskip .1cm
 $\displaystyle{  \over  \mu.  P \arr\mu
P } $  $ \hb$   
\hskip .5cm
 $\displaystyle{   P \arr\mu   P' \over   P + Q   \arr\mu
P'  } $  $ \hb$   
\hskip .5cm
 $\displaystyle{   P \arr\mu   P' \over   P | Q   \arr\mu
P' | Q } $  $ \hb$   
\hskip .3cm
  $\; \;$  $\displaystyle{ P \arr{ a}P' \hk \hk  Q
\arr{\outC a }Q'  \over  P|  Q \arr{ \tau} P'
|  Q'  }$ 
\\
\vspace{.2cm}
$\displaystyle{ P \arr{\mu}P' \over
 (\res L\!)\, P \arr{\mu} (\res L\!)\, P'} $ $ \mu, \outC\mu \not\in L$
$ \hb$
%
$\displaystyle{ P \sub {\recu X P} X \arr{\mu}P' \over
 \recu X P \arr{\mu} P'} $
\hskip .5cm  
$\displaystyle{ P \arr{\mu} P' \over
 P \;[r\!f] \arr{r\!f(\mu)} P' \;[r\!f]} $
$ \hb$ %  &
\end{center}
\caption{Structural Operational Semantics of CCS}
\label{f:LTSCCS}
\end{figure*}

Some standard notations for transitions: $\Arr\epsilon$ is the 
reflexive and transitive closure of $\arr\tau$, and 
$\Arr \mu $ is $\Arr\epsilon \arr\mu \Arr\epsilon$ (the
composition of the three relations).
Moreover,   
$P \arcap \mu P'$ holds if $P \arr\mu P'$ or $\mu =\tau \wedge P = P'$; similarly 
$P \Arcap \mu P'$ holds if $P \Arr\mu P'$ or $\mu =\tau \wedge P = P'$).
We write $P \:(\arr\mu)^n P'$ if $P$ can become $P'$ after performing
$n$ $\mu$-transitions. Finally, $P \arr\mu$ holds if there is $P'$
with $P \arr\mu P'$, and similarly for other forms of transitions.

\paragraph{Further notations}
We let $\R$, $\S$ range over binary relations, sometimes using 
 infix notation for them, \hl{e.g.} $P \,\R\, Q$ means $(P,Q) \in \R$.
We use a tilde, as in  $\til P$, to denote tuples of processes with
countably many \hl{elements.}
All \hl{relation} notations can be extended to tuples componentwise,
e.g., $\til P \,\R\, \til Q$ means $P_i \,\R\, Q_i$ for each index $i$ of the tuples $\til P$ and $\til Q$.
We use $\DSdefi$ to \hl{define} abbreviations. For instance, $P \DSdefi G $, where
$G$ is some expression, means that $P$ stands for the expression $G$.
If $\leq$ is a preorder, then  $\geq$  is its inverse (and
conversely).

>>>>

Here we recall the core theory of CCS.
We assume a possibly infinite set of names $\mathscr{L} = \{a, b,
\ldots\}$ yielding input and output labels (actions), plus a special invisible
action $\tau \notin \mathscr{L}$, and another possibly infinite
set of agent variables \hl{$\mathscr{X} = \{X, Y, \ldots\}$}.
The class of \hl{all} CCS \hl{terms (or expressions)} is then inductively
defined from $\nil$ (the terminated process) and agent variables by the operators
of \emph{prefixing} (.), \emph{parallel composition} ($|$),
\emph{summation}
(or \emph{binary choice}, $+$),
\emph{restriction} ($\nu$), \emph{relabeling} ($[\cdot]$) and
\emph{recursion} ($\texttt{rec}$):
\begin{equation*}
\begin{array}{cccl}
\mu  & := & \tau\!\!\!\! & \midd \; a  \; \midd \;  \outC a  \\
P  & := & \nil\!\!\!\! & \midd \;  \mu . P \; \midd \;  P_1 |  P_2 \; \midd  \;
P_1 + P_2 \; \midd 
(\res L)\, P
\; \midd \; P\; [\mathit{rf}]  % relabelling
\;  \midd \;  X \; \midd \; \recu X P
\end{array}
\end{equation*}
In our presentation of CCS, the restriction operator takes a set of labels $L
  \subseteq \mathscr{L}$, rather than a single one.  
The relabeling operator takes a relabeling function
  $\mathit{rf} \colon \mathscr{L} \cup \overline{\mathscr{L}} \cup
    \{\tau\} \rightarrow \mathscr{L} \cup \overline{\mathscr{L}} \cup
    \{\tau\}$, that can  handle multiple actions including $\tau$. A valid
    relabeling function $\mathit{rf}$ must however 
satisfy $\mathit{rf}(\tau) =
    \tau$ and $\forall l\in\mathscr{L} \cup \overline{\mathscr{L}}.\, \mathit{rf}(\overline{l}) =
    \overline{\mathit{rf}(l)}$ (with $\bar{\bar l} = l$ for all $l \in
    \mathscr{L}$).
We sometimes omit a trailing $\nil$, e.g., writing $a|b$ for $a.\nil |b .\nil$.
A CCS process $P$ may evolve to another one, say $P'$, under
  an action $\mu$, written by $P \arr{\mu} P'$.
The transition semantics of CCS processes is given by means of
a Labeled Transition System (LTS) expressed in Structural Operational
Semantics (SOS) rules shown in Fig.~\ref{f:LTSCCS}.
(The symmetric versions of the SOS rules for
parallel composition and binary choice are omitted.)
A CCS processes uses only \emph{guarded sums} if all occurrences of
sums are in the form $a.P + b.Q$. 
The \emph{immediate derivatives} of a
process $P$ are elements of the set $\{P' \st P \arr\mu P' \mbox{for some $\mu$}\}$.
We use $\ell$ to range over visible actions (i.e.~inputs or outputs, excluding  $\tau$)
and $\alpha, \mu$  to range over all actions.
\begin{figure*}[t]
\begin{center}
\vskip .1cm
 $\displaystyle{  \over  \mu.  P \arr\mu
P } $  $ \hb$   
\hskip .5cm
 $\displaystyle{   P \arr\mu   P' \over   P + Q   \arr\mu
P'  } $  $ \hb$   
\hskip .5cm
 $\displaystyle{   P \arr\mu   P' \over   P | Q   \arr\mu
P' | Q } $  $ \hb$   
\hskip .3cm
  $\; \;$  $\displaystyle{ P \arr{ a}P' \hk \hk  Q
\arr{\outC a }Q'  \over  P|  Q \arr{ \tau} P'
|  Q'  }$ 
\\
\vspace{.2cm}
$\displaystyle{ P \arr{\mu}P' \over
 (\res L\!)\, P \arr{\mu} (\res L\!)\, P'} $ $ \mu, \outC\mu \not\in L$
$ \hb$
%
$\displaystyle{ P \sub {\recu X P} X \arr{\mu}P' \over
 \recu X P \arr{\mu} P'} $
\hskip .5cm  
$\displaystyle{ P \arr{\mu} P' \over
 P \;[r\!f] \arr{r\!f(\mu)} P' \;[r\!f]} $
$ \hb$ %  &
\end{center}
\caption{Structural Operational Semantics of CCS}
\label{f:LTSCCS}
\end{figure*}

Some standard notations for transitions: $\Arr\epsilon$ is the 
reflexive and transitive closure of $\arr\tau$, and 
$\Arr \mu $ is $\Arr\epsilon \arr\mu \Arr\epsilon$ (the
composition of the three relations).
Moreover,   
$P \arcap \mu P'$ holds if $P \arr\mu P'$ or $\mu =\tau \wedge P = P'$; similarly 
$P \Arcap \mu P'$ holds if ($P \Arr\mu P'$ or $\mu =\tau \wedge P = P'$).
We write $P \:(\arr\mu)^n P'$ if $P$ can become $P'$ after performing
$n$ $\mu$-transitions. Finally, $P \arr\mu$ holds if there is $P'$
with $P \arr\mu P'$, and similarly for other forms of transitions.

\paragraph{Further notations}
We let $\R$, $\S$ range over binary relations, sometimes using 
 infix notation for them, \hl{e.g.} $P \,\R\, Q$ means $(P,Q) \in \R$.
We use a tilde, as in  $\til P$, to denote (finite) tuples of elements. 
% We use a tilde, as in  $\til P$, to denote tuples of processes with
% countably many \hl{elements.}
All \hl{relation} notations can be extended to tuples componentwise,
e.g., $\til P \,\R\, \til Q$ means $P_i \,\R\, Q_i$ for each index $i$ of the tuples $\til P$ and $\til Q$.
We use $\DSdefi$ to \hl{define} abbreviations. For instance, $P \DSdefi G $, where
$G$ is some expression, means that $P$ stands for the expression $G$.
If $\leq$ is a preorder, then  $\geq$  is its inverse (and
conversely).

### 2019-11-25 ic-journal/contraction.tex
Cumulative fixes with highlights
text

Intuitively, the bisimilarity contraction $\mcontrBIS$ is a preorder
such that, $P \mcontrBIS Q$ holds if $P \wb Q$ and, in addition, 
$Q$ \emph{has the possibility of being at least as efficient as $P$} (as far as
$\tau$-actions are performed).
Process $Q$, however, may be non-deterministic and may have other ways
of doing the same work, and these could be slow (i.e., involving
more $\tau$-actions than those performed by $P$).
% Thus, in contrast with expansion,  we cannot really say that `$Q$ is more efficient than
% $P$'.

@@

In the first clause $Q$ is required to match $P$'s challenge
transition with at most one transition.
This makes sure that $Q$ is capable of mimicking % verb: mimic
$P$'s work at least as efficiently as $P$. 
In contrast, the second clause of Definition~\ref{d:BisCon}, on the
challenges from $Q$, entirely ignores efficiency: it is the same
clause of  weak bisimulation~---the final derivatives are indeed required
to be related  by $\wb$, rather than by $\R$.

Bisimilarity contraction is coarser than 
the \emph{expansion relation}
$\expa$~\cite{arun1992efficiency,sangiorgi2015equations}:
\begin{definition}%[expansion]
\label{d:expa}
A process relation ${\R}$
  is an \emph{expansion} if, whenever
we have $P\RR Q$, for all $\mu$
 \begin{enumerate}
 \item   $P \arr\mu P'$ implies that there is $Q'$ with $Q \arcap \mu  Q'$
  and $P' \RR Q'$;
 \item $Q \arr\mu Q'$   implies that there is $P'$ with $P \Arr \mu P'$ and $P' \RR Q'$.
 \end{enumerate}
Two processes $P$ and $Q$ are in the \emph{bisimilarity
  expansion}, written as $P \expa Q$ ($P$ \emph{expands} $Q$),
 if $P \RR Q$ for some expansion $\R$.
 \end{definition}
The expansion preorder is widely used in proof techniques for bisimilarity and
intuitively refines bisimilarity by 
formalising the idea of ``efficiency'' between processes.
Clause (1) is the same in the both preorders, while in clause (2) expansion uses 
$P \Arr \mu P'$, rather than $P \Arcap \mu P'$; 
moreover in clause (2) of Def.~\ref{d:BisCon} the final derivatives
are simply required to be bisimilar ($P' \wb Q'$).
Intuitively, $P \expa Q$ holds if $P\wb Q$ and, in addition, \emph{$Q$
  is always at least as efficient as $P$}.

>>>>

Intuitively, the bisimilarity contraction $\mcontrBIS$ is a preorder
where $P \mcontrBIS Q$ holds if $P \wb Q$ and, in addition, 
$Q$ \emph{has the possibility of being at least as efficient as $P$} (as far as
$\tau$-actions are performed).
Process $Q$, however, may be non-deterministic and may have other ways
of doing the same work, and these could be slow (i.e., involving
more $\tau$-actions than those performed by $P$).
% Thus, in contrast with expansion,  we cannot really say that `$Q$ is more efficient than
% $P$'.

@@

In the first clause $Q$ is required to match $P$'s challenge
transition with at most one transition.
This makes sure that $Q$ is capable of mimicking % verb: mimic
$P$'s work at least as efficiently as $P$. 
In contrast, the second clause of Definition~\ref{d:BisCon}, on the
challenges from $Q$, entirely ignores efficiency: it is the same
clause of  weak bisimulation~--- the final derivatives are indeed required
to be related  by $\wb$, rather than by $\R$.

Bisimilarity contraction is coarser than 
the \emph{expansion relation}
$\expa$~\cite{arun1992efficiency,sangiorgi2015equations}:
\begin{definition}[expansion]
\label{d:expa}
A process relation ${\R}$
  is an \emph{expansion} if, whenever
we have $P\RR Q$, for all $\mu$
 \begin{enumerate}
 \item   $P \arr\mu P'$ implies that there is $Q'$ with $Q \arcap \mu  Q'$
  and $P' \RR Q'$;
 \item $Q \arr\mu Q'$   implies that there is $P'$ with $P \Arr \mu P'$ and $P' \RR Q'$.
 \end{enumerate}
Two processes $P$ and $Q$ are in the \emph{bisimilarity
  expansion}, written as $P \expa Q$ ($P$ \emph{expands} $Q$),
 if $P \RR Q$ for some expansion $\R$.
 \end{definition}
The expansion preorder is widely used in proof techniques for bisimilarity and
intuitively refines bisimilarity by 
formalising the idea of ``efficiency'' between processes.
Clause (1) is the same in the both preorders, while in clause (2) expansion uses 
$P \Arr \mu P'$, rather than $P \Arcap \mu P'$; 
moreover in clause (2) of Def.~\ref{d:BisCon} the final derivatives
are simply required to be bisimilar ($P' \wb Q'$).
Intuitively, $P \expa Q$ holds if $P\wb Q$ and, in addition, \emph{$Q$
  is always at least as efficient as $P$}.

### 2019-11-25 ic-journal/equa.tex
text

In this paper (and the formalisation work), we use the agent
variables also as \emph{equation variables}. This eliminates the need of
another type for  equations, and we can reuse the existing
variable substitution operation (c.f. SOS rule for the Recursion in
Fig.~\ref{f:LTSCCS}) for the substitution of equation variables.
For example, the result of substituting variable $X$ with $\nil$ in $a.X +
\recu X (b.X)$,  written $(a.X + \recu X (b.X)) \sub {\nil} X$, is
$a.\nil + \recu X (b.X)$ (with the part $\recu X (b.X)$
untouched). \Multivariate substitutions are written in the same syntax,
e.g. $E \sub {\til P} {\til X}$. Whenever $\til X$ is clear from the
context, we may also write $E[\til P]$ instead of $E \sub {\til P} {\til
  X}$ (and $E[P]$ for $E \sub P X$ if there is a single equation variable $X$.)

@@

The components of $\til P$ need not be
different from each other, as it must be for the variables $\til X$.
If the system has infinitely many equations, the  tuples $\til P$ and $\til X$ are infinite too.

@@

\begin{theorem}[unique solution of equations for $\rapprox$]
\label{t:Mil89s3}
Let $E_i$ be guarded and sequential with free
variables up to $\til X$, and let ${\til P} \rapprox {\til E}\{\til P /\til X\}$,
  ${\til Q} \rapprox {\til E}\{\til Q /\til X\}$. Then ${\til P} \rapprox {\til Q}$.
\end{theorem}

The version of Milner's unique-solution theorem for $\wb$ further requires
that all sums are guarded:
% \footnote{But if the CCS syntax were defined with only guarded
%   sums, i.e., $\sum_{i\in I} \mu_i.P_i$ as
%   in~\cite{sangiorgi2015equations}, this addition
%   requirement disappears automatically, and we can even say $\wb$ is indeed a
%   congruence.}:
\begin{theorem}[unique solution of equations for $\wb$]
\label{t:Mil89}
Let $E_i$ be guarded and sequential with free
variables up to $\til X$, and let ${\til P} \wb {\til E}\{\til P /\til X\}$,
  ${\til Q} \wb {\til E}\{\til Q /\til X\}$. Then ${\til P} \wb {\til Q}$.
\end{theorem}

>>>>

In this paper (and the formalisation work), we use the agent
variables also as \emph{equation variables}. This eliminates the need of
another type for  equations, and we can reuse the existing
variable substitution operation (c.f.\ the   SOS rule for the Recursion in
Fig.~\ref{f:LTSCCS}) for the substitution of equation variables.
For example, the result of substituting variable $X$ with $\nil$ in $a.X +
\recu X (b.X)$,  written $(a.X + \recu X (b.X)) \sub {\nil} X$, is
$a.\nil + \recu X (b.X)$ (with the part $\recu X (b.X)$
untouched). \Multivariate substitutions are written in the same syntax,
e.g. $E \sub {\til P} {\til X}$. Whenever $\til X$ is clear from the
context, we may also write $E[\til P]$ instead of $E \sub {\til P} {\til
  X}$ (and $E[P]$ for $E \sub P X$ if there is a single equation
variable $X$).

@@

The components of $\til P$ need not be
different from each other, as it must be for the variables $\til X$.
% If the system has infinitely many equations, the  tuples $\til P$
% and $\til X$ are infinite too.

@@

\begin{theorem}[unique solution of equations for $\rapprox$]
\label{t:Mil89s3}
Let $E_i$ be guarded and sequential with free
variables in $\til X$, and let ${\til P} \rapprox {\til E}\{\til P /\til X\}$,
  ${\til Q} \rapprox {\til E}\{\til Q /\til X\}$. Then ${\til P} \rapprox {\til Q}$.
\end{theorem}

The version of Milner's unique-solution theorem for $\wb$ further requires
that all sums are guarded:
% \footnote{But if the CCS syntax were defined with only guarded
%   sums, i.e., $\sum_{i\in I} \mu_i.P_i$ as
%   in~\cite{sangiorgi2015equations}, this addition
%   requirement disappears automatically, and we can even say $\wb$ is indeed a
%   congruence.}:
\begin{theorem}[unique solution of equations for $\wb$]
\label{t:Mil89}
Let $E_i$ be guarded and sequential with free
variables in $\til X$, and let ${\til P} \wb {\til E}\{\til P /\til X\}$,
  ${\til Q} \wb {\til E}\{\til Q /\til X\}$. Then ${\til P} \wb {\til Q}$.
\end{theorem}

### 2019-11-25 ic-journal/expansion.tex
text

\emph{Strong} bisimulation and strong bisimilarity ($\sim$)
can be obtained by replacing the weak transition $Q\Arcap\mu Q'$ in the above
definition with normal transitions $Q \arr \mu Q'$ (similar for the other case).
Weak bisimilarity is not preserved by the sum operator (except for
guarded sums), i.e.~$P_1 \wb Q_1$ and $P_2 \wb Q_2$  \hl{do} not imply
 $P_1 + P_2 \wb Q_1 + Q_2$.
For this, Milner introduced the concept of \emph{observational congruence}, also called \emph{rooted
  bisimilarity}~\cite{Gorrieri:2015jt,Sangiorgi:2011ut}:
\begin{definition}%[rooted bisimilarity]
\label{d:rootedBisimilarity}
Two processes $P$ and $Q$ are \emph{rooted bisimilar}, written as $P
\rapprox Q$, if % for all $\mu$:
%  for all $\mu\in \mathscr{L}\cup\{\tau\}$
\begin{enumerate}
 \item  $P \arr\mu P'$ implies that there is $Q'$ such that $Q
   \Arr\mu Q'$ and $P' \wb Q'$;
 \item  $Q \arr\mu Q'$ implies that there is $P'$ such that $P
   \Arr\mu P'$ and $P' \wb Q'$\enspace.
\end{enumerate}
\end{definition}
% This new sentence maintains a link with Definition 2.1
\hl{It can be proved that, $P_1 \rapprox Q_1 \wedge P_2 \rapprox Q_2 \Longrightarrow
P_1 + P_2 \rapprox Q_1 + Q_2$.}
The above definition also brings up a proof technique for proving rooted
bisimilarity from a given bisimulation. The next
lemma plays an
important role in proving some key results in this paper:
\begin{lemma}[Rooted bisimilarity by bisimulation]
\label{l:obsCongrByWeakBisim}
Given a (weak) bisimulation $\RR$, suppose two processes $P$ and $Q$
satisfy the following properties:
\begin{enumerate}
\item $P \arr\mu P'$ implies that there is $Q'$ such that $Q
   \Arr\mu Q'$ and $P' \RR Q'$;
\item $Q \arr\mu Q'$ implies that there is $P'$ such that $P
   \Arr\mu P'$ and $P' \RR Q'$.
\end{enumerate}
Then $P$ and $Q$ are rooted bisimilar, i.e.~$P \approx^c Q$.
\end{lemma}

One basic property of rooted bisimiarity is congruence,
indeed the coarsest congruence contained in bisimilarity
(See Section~\ref{s:coarsest} and \hl{\mbox{Section~\ref{ss:context}}} for more details):
\begin{theorem}
\label{t:rapproxCongruence}
$\rapprox$ is a congruence in CCS, and it is the
coarsest congruence contained in $\wb$.
\end{theorem}

>>>>

\emph{Strong} bisimulation and strong bisimilarity ($\sim$)
can be obtained by replacing the weak transition $Q\Arcap\mu Q'$ in 
clause (1) 
 with the transition $Q \arr \mu Q'$ (and similarly for the other clause).
Weak bisimilarity is not preserved by the sum operator (except for
guarded sums), i.e.~$P_1 \wb Q_1$ and $P_2 \wb Q_2$  \hl{do} not imply
 $P_1 + P_2 \wb Q_1 + Q_2$.
For this, Milner introduced the concept of \emph{observational congruence}, also called \emph{rooted
  bisimilarity}~\cite{Gorrieri:2015jt,Sangiorgi:2011ut}:
\begin{definition}%[rooted bisimilarity]
\label{d:rootedBisimilarity}
Two processes $P$ and $Q$ are \emph{rooted bisimilar}, written as $P
\rapprox Q$, if % for all $\mu$:
%  for all $\mu\in \mathscr{L}\cup\{\tau\}$
\begin{enumerate}
 \item  $P \arr\mu P'$ implies that there is $Q'$ such that $Q
   \Arr\mu Q'$ and $P' \wb Q'$;
 \item  $Q \arr\mu Q'$ implies that there is $P'$ such that $P
   \Arr\mu P'$ and $P' \wb Q'$\enspace.
\end{enumerate}
\end{definition}
% This new sentence maintains a link with Definition 2.1
\hl{It can be proved that, $P_1 \rapprox Q_1 \wedge P_2 \rapprox Q_2 \Longrightarrow
P_1 + P_2 \rapprox Q_1 + Q_2$.}
The above definition also brings up a proof technique for proving rooted
bisimilarity from a given bisimulation. The next
lemma plays an
important role in proving some key results in this paper:
\begin{lemma}[Rooted bisimilarity by bisimulation]
\label{l:obsCongrByWeakBisim}
Given a (weak) bisimulation $\RR$, suppose two processes $P$ and $Q$
satisfy the following properties:
\begin{enumerate}
\item $P \arr\mu P'$ implies that there is $Q'$ such that $Q
   \Arr\mu Q'$ and $P' \RR Q'$;
\item $Q \arr\mu Q'$ implies that there is $P'$ such that $P
   \Arr\mu P'$ and $P' \RR Q'$.
\end{enumerate}
Then $P$ and $Q$ are rooted bisimilar, i.e.~$P \approx^c Q$.
\end{lemma}

One basic property of rooted bisimiarity is congruence,
indeed it is  the coarsest congruence contained in bisimilarity
(See Section~\ref{s:coarsest} and \hl{\mbox{Section~\ref{ss:context}}} for more details):
\begin{theorem}
\label{t:rapproxCongruence}
$\rapprox$ is a congruence in CCS, and it is the
coarsest congruence contained in $\wb$.
\end{theorem}
